Global code motion of congruent instructions

### 2017-11-06 cgo/paper.tex
ssapre

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{comment}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{graphviz}
\usepackage{auto-pst-pdf}
\usepackage{etoolbox}
\usepackage{flushend}
\usepackage{needspace}
\usepackage{algpseudocode}
\usepackage{enumitem}
\usepackage[final]{pdfpages}

@@

\section{Code hoisting of congruent computations}
Examples to show CHI insertion.

The algorithm factors out the reachability of values such that multiple queries
to find reachability of values are fast. This is based on finding the ANTIC
points in the CFG which do not change during hoisting. The ANTIC points are
basically the dominance-frontiers in the inverse graph. So we introduce a data
structure (CHI nodes) to keep track of values flowing out of a basic block. We
only do this for values with multiple occurrences in the function as they are
the potentially hoistable candidates. This approach allows us to hoist
instructions to a basic block with more than two successors, as well as deal
with infinite loops in a trivial way.

@@

\begin{figure}[h!]
  \includegraphics[width=0.4\textwidth]{chi-example.pdf}
  \vspace*{-1.0cm}
\caption{I1, I2 are instructions in B, and C respectively, CHIArg1 = \{V, B,
  I1\} and CHIArg2 = \{V, C, I2\}, V is the value number for both I1, and I2}
\label{fig:chi-intro}
\end{figure}

@@

@@

\begin{verbatim}
struct CHIArg {
  // Edge destination (shows the direction of flow),
  // may not be where the I is.
  BasicBlock *Dest;
  // The instruction CHI-arg refers to.
  Instruction *I;
  // The global value number of I.
  GVNType VN;
};

\end{verbatim}

Essentially, CHIArg1 = \{V, B, I1\} and CHIArg2 = \{V, C, I2\}

For sinking we use the similar approach as \cite{ssapre} to track values flowing
into a basic block.

>>>>

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption
\usepackage{comment}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{graphviz}
\usepackage{auto-pst-pdf}
\usepackage{etoolbox}
\usepackage{flushend}
\usepackage{needspace}
\usepackage{algpseudocode}
\usepackage{enumitem}
\usepackage[final]{pdfpages}

@@

\section{A fastt Code motion of congruent computations}
The algorithm factors out the reachability of values such that multiple queries
to find reachability of values are fast. This is based on finding the dominance
frontiers and the post-dominance frontiers of the CFG. These points, being
structural properties of a graph, do not change during code-motion.

A post-dominance frontier is the basic block where anticipability (ANTIC) of an
instruction (I) is checked because it the point on which the basic block
containing I is control dependent.

So we introduce a data structure (CHI nodes) to keep track of values flowing out
of a basic block. We only do this for values with multiple occurrences in the
function as they are the potentially hoistable candidates. This approach allows
us to hoist instructions to a basic block with more than two successors, as well
as deal with infinite loops in a trivial way.

@@

@@

\begin{figure}
  \includegraphics[width=0.4\textwidth]{chi-example.pdf}
  \vspace*{-1.0cm}
\caption{I1, I2 are instructions in B, and C respectively, CHIArg1 = \{V, B,
  I1\} and CHIArg2 = \{V, C, I2\}, V is the value number for both I1, and I2}
\label{fig:chi-intro}
\end{figure}

@@

\begin{verbatim}
struct CHIArg {
  // Edge destination (shows the direction of flow),
  // may not be where the I is.
  BasicBlock *Dest;
  // The instruction CHI-arg refers to.
  Instruction *I;
  // The global value number of I.
  GVNType VN;
};
\end{verbatim}

Essentially, CHIArg1 = \{V, B, I1\} and CHIArg2 = \{V, C, I2\}. For sinking we
use the similar approach as \cite{ssapre} to track values flowing into a basic
block.

When SSAPRE inserts PHI (expression PHIs), it does so at places where
(potentially) multiple expressions may merge. If a PHI has a missing entry, that
means the expression is partially available.  The concept may work for hoisting
expression to help factor out anticipability by working on inverted graph. If we
insert a CHI, to a basic block with multiple successors, as instructions are
hoisted upwards to the post-dominance frontier.  And we start walking the CFG to
figure out if any CHI has a missing entry, that means the expression is not
anticipable in the basic block having that outgoing PHI.

\begin{itemize}
\item Sort VNs according to their rank, and start with lowest ranked VN
\item Take a VN and for each instruction with same VN
\item  - Find the dominance frontier in the inverse graph (PDF)
\item  - Insert the chi-node at PDF
\item Remove the chi-nodes with missing entries
\item Remove values from CHI-nodes which do not truly flow out, e.g.,
\item  modified along the path.
\item Collect the remaining values that are still anticipable
\end{itemize}
